---
title: "Lab 7: Functions + Fish"
author: "Kallan Torell"
format: html
editor: source
embed-resources: true
execute: 
  code-tools: true
  eval: true
  echo: true
---

The goal of this lab is learn more about exploring missing data and writing
modular code.

```{r}
#| label: setup

library(tidyverse)
fish <- read.csv("BlackfootFish.csv")
glimpse(fish)
```

## The Data

This lab's data concerns mark-recapture data on four species of trout from the
Blackfoot River outside of Helena, Montana. These four species are
**rainbow trout (RBT)**, **westslope cutthroat trout (WCT)**, **bull trout**,
and **brown trout**.

Mark-recapture is a common method used by ecologists to estimate a population's
size when it is impossible to conduct a census (count every animal). This method
works by *tagging* animals with a tracking device so that scientists can track
their movement and presence.

## Data Exploration

The measurements of each captured fish were taken by a biologist on a raft in
the river. The lack of a laboratory setting opens the door to the possibility of
measurement errors.

**1. Let's look for missing values in the dataset. Output ONE table that answers BOTH of the following questions:**

+ **How many observations have missing values?**
+ **What variable(s) have missing values present?**

::: callout-tip
# You should use `across()`!
:::

```{r}
#| label: find-missing-values

fish |>
  summarize(
    across(
      .cols  = everything(),
      .fns   = ~ sum(is.na(.)),
      .names = "missing_{col}"
    )
  ) |>
  pivot_longer(
    cols      = everything(),
    names_to  = "variable",
    values_to = "n_missing"
  ) |>
  mutate(
    variable = str_remove(variable, "^missing_")
  ) |>
  filter(n_missing > 0)

# https://www.rdocumentation.org/packages/tidyselect/versions/1.2.1/topics/everything
# been using everything() and i dont recall using it in class since I missed last week but it selects all remaining variables after other operations have been applied, like in this case pivot_longer() or other cases could be select() or across()
```



**2. Use a pivot to transform your summary table from Question 1 into a table that is easier to read.** 

```{r}
#| label: pivot-summary-table

fish |>
  summarize(
    across(
      .cols  = everything(),
      .fns   = ~ sum(is.na(.)),
      .names = "missing_{col}"
    )
  ) |>
  pivot_longer(
    cols      = everything(),
    names_to  = "variable",
    values_to = "count"
  ) |>
  mutate(variable = str_remove(variable, "^missing_")) |>
  pivot_wider(
    names_from   = variable,
    values_from  = count,
    values_fill  = 0,
    names_prefix = "missing_"
  )

```


**3. Create ONE thoughtful visualization that explores the frequency of missing values across the different years, sections, and trips.**

```{r}
#| label: visual-of-missing-values-over-time

library(ggplot2)

missing_weight_by_year_section_trip <-
  fish |>
  filter(is.na(weight)) |>
  count(year, section, trip, name = "n_missing")


ggplot(
  missing_weight_by_year_section_trip,
  aes(x = factor(trip), y = n_missing, fill = section)
) +
  geom_col(position = "dodge") +
  facet_wrap(~ year, scales = "free_y") +
  labs(
    title    = "Missing Weight Values by Trip and Section",
    subtitle = "Faceted by year; only rows with missing weight are included",
    x        = "Trip",
    y        = "Count of missing weights",
    fill     = "Section"
  )

#https://ggplot2.tidyverse.org/reference/position_dodge.html
# posotion dodge for geom_col isnt in class i believe but for a plot like this it just makes it easier to read, so for this plot, the bars will be placed side by side and not vertically stacked, I thought it would be easier to read
  
```

```{r}
missing_by_year <-
  fish |>
  filter(is.na(weight)) |>
  count(year, section, name = "n_missing")

ggplot(missing_by_year, aes(x = year, y = n_missing, color = section, group = section)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Missing weight counts by year",
    x = "Year", y = "Count of missing weights", color = "Section"
  )

```



## Rescaling the Data

If I wanted to rescale every quantitative variable in my dataset so that they
only have values between 0 and 1, I could use this code:

```{r}
#| echo: true
#| eval: false

fish <- fish |> 
  mutate(length = (length - min(length, na.rm = TRUE)) / 
           (max(length, na.rm = TRUE) - min(length, na.rm = TRUE)), 
         weight = (weight - min(weight, na.rm = TRUE)) / 
           (max(weight, na.rm = TRUE) - min(length, na.rm = TRUE)))
```


**4. Transform the repeated process above into a `rescale_01()` function. Your function should...**

+ **... take a single vector as input.**
+ **... return the rescaled vector.**

```{r}
#| label: write-rescale-function

rescale_01 <- function(x, na.rm = TRUE) {
  if (!is.numeric(x)) stop("`x` must be numeric.")
  if (length(x) <= 1) stop("`x` must have length > 1.")

  if (all(is.na(x))) return(x)
  r <- range(x, na.rm = na.rm)
  rng <- r[2] - r[1]
  if (rng == 0) return(ifelse(is.na(x), NA_real_, 0))
  (x - r[1]) / rng
}


```

::: callout-tip
# Efficiency 

Think about the efficiency of the function you wrote. Are you calling the
**same** function multiple times? You might want to look into the `range()` 
function. 
:::

**5. Let's incorporate some input validation into your function. Modify your previous code so that the function stops if ...**

+ **... the input vector is not numeric.**
+ **... the length of the input vector is not greater than 1.**

::: callout-tip
# Modify Previous Code

Do not create a new code chunk here -- simply add these stops to your function
above!
:::

## Test Your Function

**6. Run the code below to test your function. Verify that the maximum of your rescaled vector is 1 and the minimum is 0!**

```{r}
#| label: verify-rescale-function

x <- c(1:25, NA)

rescaled <- rescale_01(x)
min(rescaled, na.rm = TRUE)
max(rescaled, na.rm = TRUE)
```

Next, let's test the function on the `length` column of the `BlackfootFish` data.

**7. The code below makes a histogram of the original values of `length`. Add a plot of the rescaled values of `length`. Output your plots side-by-side, so the reader can confirm the only aspect that has changed is the scale.**

::: callout-warning
This will require you to call your `rescale_01()` function within a `mutate()`
statement in order to create a `length_scaled` variable.
:::

```{r}
#| label: compare-original-with-rescaled-lengths



# Code for Q7 plot.

fish |>
  mutate(length_scaled = rescale_01(length)) |>
  select(length, length_scaled) |>
  pivot_longer(everything(), names_to = "measure", values_to = "value") |>
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30, na.rm = TRUE) +
  facet_wrap(
    ~ measure,
    scales = "free_x",
    labeller = as_labeller(c(
      length = "Original length (mm)",
      length_scaled = "Rescaled length (0–1)"
    ))
    #https://r-charts.com/ggplot2/facets/ 
    # I used this online source just for the many arguments of facet_wrap and to make sure I was using it right
  ) +
  labs(
    title = "Fish length: original vs rescaled",
    x = NULL, y = "Count"
  )




```

::: callout-tip
1. Set the y-axis limits for both plots to go from 0 to 4000 to allow for direct comparison across plots.

2. Pay attention to `binwidth`!

3. Use a Quarto code chunk option to put the plots side-by-side.
:::