---
title: "STAT 331 Portfolio"
author: "Your Name Goes Here"
format: html 
embed-resources: true
layout: margin-left
editor: visual
execute: 
  eval: false
  echo: false
---

[**My Grade:**]{.underline} I believe my grade equivalent to course work evidenced below to be an \_\_.

[**Learning Objective Evidence:**]{.underline} In the code chunks below, provide code from Lab or Challenge assignments where you believe you have demonstrated proficiency with the specified learning target. Be sure to specify **where** the code came from (e.g., Lab 4 Question 2).

## Working with Data

**WD-1: I can import data from a *variety* of formats (e.g., csv, xlsx, txt, etc.).**

-   `csv` Example 1

```{r}
#| label: wd-1-csv-1
#| label: Lab 4 Problem 0

#| label: load-data

childcare_costs <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-05-09/childcare_costs.csv')

counties <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-05-09/counties.csv')

tax_rev <- read_csv('https://atheobold.github.io/groupworthy-data-science/labs/instructions/data/ca_tax_revenue.csv')

head(childcare_costs)
```

-   `csv` Example 2

```{r}
#| label: wd-1-csv-2
#| label: Lab 2 Problem 1

surveys <- read_csv("surveys.csv")
head(surveys)
glimpse(surveys)

```

-   `xlsx`

```{r}
#| label: wd-1-xlsx
#| label: Challenge 4: Childcare Prices Problem 0

my_data <- read_excel("C:/Users/Kallan/Downloads/NDCP2022.xlsx")

head(my_data)


```

**WD-2: I can select necessary columns from a dataset.**

-   Example selecting specified columns

```{r}
#| label: wd-2-ex-1
#| Lab 2 Problem 2

ca_childcare <- counties |>
  filter(state_abbreviation == "CA" | state_name == "California") |>
  select(state_name, state_abbreviation, county_name, county_fips_code) |>
  inner_join(
    childcare_costs |>
      mutate(county_fips = county_fips_code),
    by = "county_fips_code"
  )

glimpse(ca_childcare)

```

-   Example removing specified columns

```{r}
#| label: wd-2-ex-2
#| Lab 2 Problem 2

ca_childcare <- counties |>
  filter(state_abbreviation == "CA" | state_name == "California") |>
  select(state_name, state_abbreviation, county_name, county_fips_code) |>
  inner_join(
    childcare_costs |>
      mutate(county_fips = county_fips_code),
    by = "county_fips_code"
  )

glimpse(ca_childcare)

```

-   Example selecting columns based on logical values (e.g., `starts_with()`, `ends_with()`, `contains()`, `where()`)

```{r}
#| label: wd-2-ex-3

```

**WD-3: I can filter rows from a dataframe for a *variety* of data types (e.g., numeric, integer, character, factor, date).**

-   Numeric Example 1

```{r}
#| label: wd-3-numeric-ex-1
#| lab 4 problem 6

lowest_price_2018 <- ca_childcare |>
  filter(study_year == 2018) |>
  group_by(census_region) |>
  summarise(median_infant_price = median(mc_infant, na.rm = TRUE)) |>
  slice_min(median_infant_price)

```

-   Numeric Example 2

```{r}
#| label: wd-3-numeric-ex-1
#| lab 5 problem gym investigation

checkins <- get_fit_now_check_in |>
  filter(check_in_date == "20180109") |>
  select(membership_id)
#https://dplyr.tidyverse.org/reference/select.html
#just needed syntax help and filtering help
gym_checkins <- members |>
  inner_join(checkins, by = c("id" = "membership_id")) |>
  select(person_id, name)

```

-   Character Example 1 (any context)

```{r}
#| label: wd-3-character
#| lab 4 problem 2
ca_childcare <- counties |>
  filter(state_abbreviation == "CA" | state_name == "California") |>
  select(state_name, state_abbreviation, county_name, county_fips_code) |>
  inner_join(
    childcare_costs |>
      mutate(county_fips = county_fips_code),
    by = "county_fips_code"
  )

```

-   Character Example 2 (example must use functions from **stringr**)

```{r}
#| label: wd-3-string
#| lab 4 problem final suspect

suspects <- gym_checkins |>
  left_join(person |> select(-name), by = c("person_id" = "id")) |>
  left_join(drivers_license, by = c("license_id" = "id")) |>
  transmute(person_id, name, plate_number, gender)
# https://dplyr.tidyverse.org/reference/transmute.html
# wanted help with the syntax on column management using transmute
final_suspect <- suspects |>
  filter(str_detect(plate_number, "H42W"), gender == "male")

```

-   Date (example must use functions from **lubridate**)

```{r}
#| label: wd-3-date

```

**WD-4: I can modify existing variables and create new variables in a dataframe for a *variety* of data types (e.g., numeric, integer, character, factor, date).**

-   Numeric Example 1

```{r}
#| label: wd-4-numeric-ex-1

```

-   Numeric Example 2

```{r}
#| label: wd-4-numeric-ex-2

```

-   Factor Example 1 (renaming levels)

```{r}
#| label: wd-4-factor-ex-1

```

-   Factor Example 2 (reordering levels)

```{r}
#| label: wd-4-factor-ex-2

```

-   Character (example must use functions from **stringr**)

```{r}
#| label: wd-4-string

```

-   Date (example must use functions from **lubridate**)

```{r}
#| label: wd-4-date

```

**WD-5: I can use mutating joins to combine multiple dataframes.**

-   `left_join()` Example 1

```{r}
#| label: wd-5-left-ex-1
#| lab 4 problem 3
ca_childcare <- counties |>
  filter(state_abbreviation == "CA") |>
  select(state_name, state_abbreviation, county_name, county_fips_code) |>
  inner_join(childcare_costs, by = "county_fips_code") |>
  left_join(
    tax_rev |>
      rename(county_name = entity_name) |>
      select(county_name, year, total_property_taxes, sales_and_use_taxes),
    by = c("county_name" = "county_name", "study_year" = "year")
  )
```

-   `right_join()` Example 1

```{r}
#| label: wd-5-right

```

-   `left_join()` **or** `right_join()` Example 2

```{r}
#| label: wd-5-left-right-ex-2
#| lab 5 problem gym investigation
suspects <- gym_checkins |>
  left_join(person |> select(-name), by = c("person_id" = "id")) |>
  left_join(drivers_license, by = c("license_id" = "id")) |>
  transmute(person_id, name, plate_number, gender)
# https://dplyr.tidyverse.org/reference/transmute.html
# wanted help with the syntax on column management using transmute
final_suspect <- suspects |>
  filter(str_detect(plate_number, "H42W"), gender == "male")

final_suspect
```

-   `inner_join()` Example 1

```{r}
#| label: wd-5-inner-ex-1
#| lab 4 problem 3
ca_childcare <- counties |>
  filter(state_abbreviation == "CA") |>
  select(state_name, state_abbreviation, county_name, county_fips_code) |>
  inner_join(childcare_costs, by = "county_fips_code") |>
  left_join(
    tax_rev |>
      rename(county_name = entity_name) |>
      select(county_name, year, total_property_taxes, sales_and_use_taxes),
    by = c("county_name" = "county_name", "study_year" = "year")
  )
```

-   `inner_join()` Example 2

```{r}
#| label: wd-5-inner-ex-2
#| lab 5 problem gym investigation
checkins <- get_fit_now_check_in |>
  filter(check_in_date == "20180109") |>
  select(membership_id)
#https://dplyr.tidyverse.org/reference/select.html
#just needed syntax help and filtering help
gym_checkins <- members |>
  inner_join(checkins, by = c("id" = "membership_id")) |>
  select(person_id, name)

gym_checkins

```

**WD-6: I can use filtering joins to filter rows from a dataframe.**

-   `semi_join()`

```{r}
#| label: wd-6-semi

```

-   `anti_join()`

```{r}
#| label: wd-6-anti

```

**WD-7: I can pivot dataframes from long to wide and visa versa**

-   `pivot_longer()`

```{r}
#| label: wd-7-long
#| lab 4 problem 7
dataplot <- ca_childcare |>
  select(census_region, study_year, mc_infant, mc_toddler, mc_preschool) |>
  pivot_longer(mc_infant:mc_preschool, names_to = "age", values_to = "price") |>
  mutate(
    age = recode(age,
                 mc_infant = "Infant",
                 mc_toddler = "Toddler",
                 mc_preschool = "Preschool"),
    age = factor(age, levels = c("Infant", "Toddler", "Preschool"))
  )
```

-   `pivot_wider()`

```{r}
#| label: wd-7-wide

```

## Reproducibility

**R-1: I can create professional looking, reproducible analyses using RStudio projects, Quarto documents, and the here package.**

The following assignments satisfy the above criteria:

-   Example 1
-   Example 2
-   Example 3
-   Example 4
-   Example 5

**R-2: I can write well documented and tidy code.**

-   Example of **ggplot2** plotting

```{r}
#| label: r-2-1
#| lab 2 problem 4

ggplot(data = surveys, 
       mapping = aes(x = hindfoot_length, y = weight)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~ species) +
  labs(
    x = "Hindfoot Length (mm)",
    y = "",
    title = "Scatterplot of Weight vs Hindfoot Length by Species",
    subtitle = "Weight (g) on y-axis"
  ) 

```

-   Example of **dplyr** pipeline

```{r}
#| label: r-2-2
#| lab 4 problem 6

lowest_price_2018 <- ca_childcare |>
  filter(study_year == 2018) |>
  group_by(census_region) |>
  summarise(median_infant_price = median(mc_infant, na.rm = TRUE)) |>
  slice_min(median_infant_price)


```

-   Example of function formatting

```{r}
#| label: r-2-3

```

**R-3: I can write robust programs that are resistant to changes in inputs.**

-   Example (any context)

```{r}
#| label: r-3-example

```

-   Example (function stops)

```{r}
#| label: r-3-function-stops

```

## Data Visualization & Summarization

**DVS-1: I can create visualizations for a *variety* of variable types (e.g., numeric, character, factor, date)**

-   At least two numeric variables

```{r}
#| label: dvs-1-num
#| Lab 2 problem 4

ggplot(data = surveys, 
       mapping = aes(x = hindfoot_length, y = weight)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~ species) +
  labs(
    x = "Hindfoot Length (mm)",
    y = "",
    title = "Scatterplot of Weight vs Hindfoot Length by Species",
    subtitle = "Weight (g) on y-axis"
  ) 

```

-   At least one numeric variable and one categorical variable

```{r}
#| label: dvs-2-num-cat
#| lab 2 problem 10

ggplot(data = surveys, aes(x = species, y = weight)) +
  geom_boxplot(alpha = 0.5) +
  labs(
    x = "Species",
    y = "Weight (g)",
    title = "Distribution of Weight by Species"
  ) + 
  geom_jitter(color = 'burlywood4') + 
  theme(axis.text.x = element_text(angle = 45))

```

-   At least two categorical variables

```{r}
#| label: dvs-2-cat
#| lab 4 problem 7

dataplot <- ca_childcare |>
  select(census_region, study_year, mc_infant, mc_toddler, mc_preschool) |>
  pivot_longer(mc_infant:mc_preschool, names_to = "age", values_to = "price") |>
  mutate(
    age = recode(age,
                 mc_infant = "Infant",
                 mc_toddler = "Toddler",
                 mc_preschool = "Preschool"),
    age = factor(age, levels = c("Infant", "Toddler", "Preschool"))
  )

ggplot(dataplot, aes(study_year, price, color = census_region)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "loess", se = FALSE) +
  facet_wrap(~ age, nrow = 1) +
  scale_x_continuous(breaks = seq(2008, 2018, 2)) +
  labs(
    title = "Weekly Median Price for Center-Based Childcare ($)",
    x = "Study Year", y = "Median Weekly Price ($)",
    color = "California Region"
  )

```

-   Dates (time series plot)

```{r}
#| label: dvs-2-date
#| lab 4 problem 8

ggplot(ca_childcare, aes(x = mhi_2018, y = mc_infant)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Income vs. Infant Center-Based Weekly Price (CA)",
    x = "Median household income (2018 Dollars)",
    y = "Median weekly price (Dollars)"
  )


```

**DVS-2: I use plot modifications to make my visualization clear to the reader.**

-   I can modify my plot theme to be more readable

```{r}
#| label: dvs-2-ex-1
#| lab 2 problem 16

ggplot(data = surveys, aes(x = weight, y = species)) +
  geom_boxplot(alpha = 0.5) +
  labs(
    x = "Species",
    y = "Weight (g)",
    title = "Distribution of Weight by Species"
  ) + 
  geom_jitter(color = 'steelblue') + 
  theme(axis.text.x = element_text(angle = 45))

```

-   I can modify my colors to be accessible to anyone's eyes

```{r}
#| label: dvs-2-ex-2
#| lab 4 problem 7

dataplot <- ca_childcare |>
  select(census_region, study_year, mc_infant, mc_toddler, mc_preschool) |>
  pivot_longer(mc_infant:mc_preschool, names_to = "age", values_to = "price") |>
  mutate(
    age = recode(age,
                 mc_infant = "Infant",
                 mc_toddler = "Toddler",
                 mc_preschool = "Preschool"),
    age = factor(age, levels = c("Infant", "Toddler", "Preschool"))
  )

ggplot(dataplot, aes(study_year, price, color = census_region)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "loess", se = FALSE) +
  facet_wrap(~ age, nrow = 1) +
  scale_x_continuous(breaks = seq(2008, 2018, 2)) +
  labs(
    title = "Weekly Median Price for Center-Based Childcare ($)",
    x = "Study Year", y = "Median Weekly Price ($)",
    color = "California Region"
  )

```

-   I can modify my plot titles to clearly communicate the data context

```{r}
#| label: dvs-2-ex-3
#| lab 2 problem 4

ggplot(data = surveys, 
       mapping = aes(x = hindfoot_length, y = weight)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~ species) +
  labs(
    x = "Hindfoot Length (mm)",
    y = "",
    title = "Scatterplot of Weight vs Hindfoot Length by Species",
    subtitle = "Weight (g) on y-axis"
  ) 

```

-   I can modify the text in my plot to be more readable

```{r}
#| label: dvs-2-ex-4
#| lab 2 problem 16

ggplot(data = surveys, aes(x = weight, y = species)) +
  geom_boxplot(alpha = 0.5) +
  labs(
    x = "Species",
    y = "Weight (g)",
    title = "Distribution of Weight by Species"
  ) + 
  geom_jitter(color = 'steelblue') + 
  theme(axis.text.x = element_text(angle = 45))
```

-   I can reorder my legend to align with the colors in my plot

```{r}
#| label: dvs-2-ex-5

```

**DVS-3: I show creativity in my visualizations**

-   I can use non-standard colors (Example 1)

```{r}
#| label: dvs-3-1-ex-1
#| lab 2 problem 3

ggplot(data = surveys,
       mapping = aes(x = weight, y = hindfoot_length)) +
  geom_point(color = 'cadetblue')

```

-   I can use non-standard colors (Example 2)

```{r}
#| label: dvs-3-1-ex-2
#| lab 2 problem 10

ggplot(data = surveys, aes(x = species, y = weight)) +
  geom_boxplot(alpha = 0.5) +
  labs(
    x = "Species",
    y = "Weight (g)",
    title = "Distribution of Weight by Species"
  ) + 
  geom_jitter(color = 'burlywood4') + 
  theme(axis.text.x = element_text(angle = 45))

```

-   I can use annotations (e.g., `geom_text()`)

```{r}
#| label: dvs-3-2

```

-   I can choose creative geometries (e.g., `geom_segment()`, `geom_ribbon)()`)

```{r}
#| label: dvs-3-3

```

**DVS-4: I can calculate numerical summaries of variables.**

-   Example using `summarize()`

```{r}
#| label: dvs-4-summarize
#| lab 4 problem 6

lowest_price_2018 <- ca_childcare |>
  filter(study_year == 2018) |>
  group_by(census_region) |>
  summarise(median_infant_price = median(mc_infant, na.rm = TRUE)) |>
  slice_min(median_infant_price)

```

-   Example using `across()`

```{r}
#| label: dvs-4-across

```

**DVS-5: I can find summaries of variables across multiple groups.**

-   Example 1

```{r}
#| label: dvs-5-1

```

-   Example 2

```{r}
#| label: dvs-5-2

```

**DVS-6: I can create tables which make my summaries clear to the reader.**

-   I can modify my column names to clearly communicate the data context

```{r}
#| label: dvs-6-ex-1

```

-   I can modify the text in my table to be more readable (e.g., bold face for column headers)

```{r}
#| label: dvs-6-ex-2

```

-   I can arrange my table to have an intuitive ordering

```{r}
#| label: dvs-6-ex-3

```

**DVS-7: I show creativity in my tables.**

-   I can use non-default colors

```{r}
#| label: dvs-7-ex-1

```

-   I can modify the layout of my table to be more readable (e.g., `pivot_longer()` or `pivot_wider()`)

```{r}
#| label: dvs-7-ex-2

```

## Program Efficiency

**PE-1: I can write concise code which does not repeat itself.**

-   using a single function call with multiple inputs (rather than multiple function calls)

```{r}
#| label: pe-1-one-call

```

-   using `across()`

```{r}
#| label: pe-1-across

```

-   using functions from the `map()` family

```{r}
#| label: pe-1-map-1

```

**PE-2: I can write functions to reduce repetition in my code.**

-   Example 1: Function that operates on vectors

```{r}
#| label: pe-2-1

```

-   Example 2: Function that operates on data frames

```{r}
#| label: pe-2-2

```

-   Example 3: Function that operates on vectors *or* data frames

```{r}
#| label: pe-2-3

```

**PE-3:I can use iteration to reduce repetition in my code.**

-   using `across()`

```{r}
#| label: pe-3-across

```

-   using a `map()` function with **one** input (e.g., `map()`, `map_chr()`, `map_dbl()`, etc.)

```{r}
#| label: pe-3-map-1

```

-   using a `map()` function with **more than one** input (e.g., `map_2()` or `pmap()`)

```{r}
#| label: pe-3-map-2

```

**PE-4: I can use modern tools when carrying out my analysis.**

-   I can use functions which are not superseded or deprecated

```{r}
#| label: pe-4-1

```

-   I can connect a data wrangling pipeline into a `ggplot()`

```{r}
#| label: pe-4-2

```

## Data Simulation & Statisical Models

**DSSM-1: I can simulate data from a *variety* of probability models.**

-   Example 1

```{r}
#| label: dsm-1-1

```

-   Example 2

```{r}
#| label: dsm-1-2

```

**DSSM-2: I can conduct common statistical analyses in R.**

-   Example 1

```{r}
#| label: dsm-2-1

```

-   Example 2

```{r}
#| label: dsm-2-2

```

-   Example 3

```{r}
#| label: dsm-2-3

```

## Revising My Thinking

<!-- How did you revise your thinking throughout the course? How did you revise your thinking on the code examples you have provided in your portfolio? -->

Throughout finding examples in my code, I realized there are certain categories that I seem to skip over or not utilize as much. Whether it be certain types of joins, not making plots with ggplot using only categorical variables, or just clustered functions, I need to diversify my coding approaches in the second half of the quarter.

<!-- For the revisions included in your Portfolio, to help me understand the nature of your revisions, please denote somehow the feedback I provided you (e.g., boldface, italics, colored text) before your revisions. -->

## Extending My Thinking

<!-- How did you extended your thinking throughout the course? How did you extend your thinking on the code examples you have provided in your portfolio? -->

At the start of this course, I approached all problems the same way I would a python data analytics problem. I realized quickly that I cannot approach R and Python in the same way, and needed to learn how to approach problems in R studio in a different way, which will help me expand this course. Many of the problems in the portfolio so far I have approached as python problems, and to diversify my thinking more, i need to learn to approach R seperately.

## Peer Support & Collaboration

<!-- Include an image or a description of feedback you gave that you are proud of (either in a peer review or in Discord). -->

<!-- Include a description of how you grew as a collaborator through the weekly pair programming activities.   -->

As a peer collaborator, I struggled in the first few PA's because I did not allow my collaborators to think through problems themselves. I often jumped in too fast to help or filled in code segments for them without discussing it. However, after a few weeks, I felt much more at ease trusting my collaborating partners, and was able to sit back and let them think when they were the coder, which I believe helped them to allow me to think through things myself and grow as a coder as well.
